The performance of presence-based and process-based species distribution models under realistic conditions

### 2020-02-13 ms/ms_latex/ms_sdmComp.tex
revised

% Range boundaries
IPMs predict $\lambda$ in each cell of the landscape, and consequently directly predict a $\lambda$-based range boundary. Similarly, the simulation-based CA models predict abundance, and thus directly predict a $N$-based range boundary. For Maxent, the range boundary definition is less explicit. The geo-located presences used as input imply an abundance-based range, but predicted distributions are frequently constrained by a data-based threshold (e.g., relative suitability values that encapsulate 95\% of the observed data) to exclude both erroneous data and sink populations, which implies an effort to predict a $\lambda$-based range. There may or may not be a big difference between the two range definitions, but it will depend on the species. For example, the magnitude of the difference should vary with the number or distribution of marginal cells (lots of marginal cells mean large areas with middling occupancy probabilities), the typical annual stochasticity in the population abundances (more stochasticity means more discrepancies are likely), and dispersal limitation (high dispersal or a long introduction history means fewer areas that have not been reached, but also possibly a larger proportion of sink populations).

>>>>

% Range boundaries
IPMs predict $\lambda$ in each cell of the landscape, and consequently directly predict the $\lambda$-based range boundary. Similarly, the simulation-based CA models predict abundance, thus directly predicting the $N$-based range boundary. For Maxent, the range boundary definition is less explicit. The geo-located presences used as input imply an abundance-based range, but predicted distributions are frequently constrained by a data-based threshold (e.g., relative suitability values that encapsulate 95\% of the observed data) to exclude both erroneous data and sink populations \citep{Elith2009a,Allen2016}, which implies an intent to predict a $\lambda$-based range. The practical difference between the range definitions will depend on the biology of the species. For example, the magnitude of the difference should vary with the number or distribution of marginal cells, where large areas with marginal cells may result in different patterns of occupancy. Similarly, species with high annual stochasticity in population abundances are more likely to exhibit patchier distributions due to chance local extinctions \citep{Morris2003}. Alternatively, species with high dispersal, or invasive species with a longer introduction history, are less likely to be dispersal limited, but may instead have a larger proportion of non-viable sink populations. Selection of process-based SDMs requires clearly specifying the range boundary definition in a way that occurrence-based SDMs do not.
